Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-07-06 analysis/paper/paper.Rmd
edit chronology section based on R1 comments

Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade network could go beyond the temporal boundaries and last longer. This is because the use of foreign goods in many Indigenous societies often go through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012]. Also, we acknowledged that a fine-grained identification of decades of the European presences from the archaeological record is challenging in our research based on some reasons. The current radiocarbon ages that are unable to provide a high resolution of time scale, the circulation of foreign goods might start earlier before the European presences, and the use of the foreign goods in Indigenous societies could last longer after the official end of European colonization [@Mitchell2010]. It is more suitable to use pericolonial approaches that consider indirect and wide-ranging effects of the colonial process such as trading activities [@Trabert2018; @Scheiber2012].

Based on these six units and copies of original excavation records, we have made some refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods. The methods we used to devise the new chronology are described in the Supplementary Online Materials.

The archaeological indicators of the start of the European influence at Kiwulan are the appearance of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages. The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Dizon2016; @Hsieh1995]. Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007]. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in north Taiwan. In any case, the high volume of ceramics transported by Europeans, and their high mobility in the shipping trade played an important role in introducing foreign jars to Taiwan.

>>>>

Despite the relatively short period of the European colonization in Taiwan (38-years-long Dutch occupation and 16-years-long Spanish occupation), the indirect impacts of the Europeans through trade network could go beyond the temporal boundaries and last longer. This is because the use of foreign goods in many Indigenous societies often go through the processes of negotiations, resistance, or transformation that a static time frame may not capture the processes well [@Scheiber2012]. Also, we acknowledged that a fine-grained identification of decades of the European presence from the archaeological record is challenging for this study based on three reasons. The current radiocarbon ages that are unable to provide a high resolution of time breaks, the circulation of foreign goods might start earlier before the official European presence in Taiwan, and the use of the foreign goods in Indigenous societies could continue after the end of the European colonization [@Mitchell2010]. However, those limitations can be minimized with pericolonial approaches that emphasize the wide-ranging effects of the colonial process, allowing discussion of indirect impacts on Indigenous societies with a more practical and dynamic framework not necessarily defined by a fixed colonial period with a start and an end time [@Trabert2018; @Scheiber2012].

We acknowledged the approach to study a continuum of changes of archaeological records over time to avoid the pitfall of temporal divides [@Scheiber2012]; however, we also admitted the convenience of using separated time periods for quantitative comparison between temporal groups of imported ornaments. To mitigate the conflict, we separated the ornaments based on the archaeological context that can reflect the acceptance of diagnostic imported goods by their relatively high frequencies as an indicator of indirect colonial impacts. That is, we separated the time periods to before and after a certain foreign presence using the complete archaeological context that contains higher frequencies of time indicators instead of their first appearance in a given layer. This method is reasonable because it allows not only the flexibility of imported goods entering Indigenous societies depending on the agency of local people, but also the application of quantitative analysis to temporal and spatial changes of imported ornaments.

Based on the copies of original excavation records, we have made refinements of the original chronology to assign contexts into the pre-European, European, and the Chinese periods with six steps of assigning and evaluations. First, we determined the context with indirect colonial impacts including the Europeans and the Chinese according to the higher frequency of time indicators. Second, we examined the description of color and texture changes in strategrphic data to identify the end of the context. Third, we checked whether there are radiocarbon ages from the context and if there are any conflicts or ambiguities to decide the usefulness of the ages. Fourth, we evaluated if there are any post-depositional issues to exclude problematic data. Fifth, we cross evaluated all fourty squares to make sure a consistent stratigraphic context across them. Sixth, we cross evaluated our chronology by checking previous studies of assigning chronology to make a final decision [@Hsieh2009; @Wang2011]. We also provide Supplementary Online Materials that include the details of stratigraphic data, the high frequency of time indicators, and the decision making during the assigning process.

The archaeological indicators of the indirect impact of the Europeans at Kiwulan are the higher frequency of light grey glazed jars, known as "An-ping" jars in China and Taiwan, and large dark brown glazed stoneware jars that were introduced to Taiwan during the early 17th century. Large dark brown glazed stoneware jars may have been made in Southeast Asia, but are frequently found in European shipwrecks from this period as vessels for transporting water, wine or other liquids on long voyages. The earliest evidence of light gray glazed jars in this region has been found among the cargo of the Spanish shipwreck _San Diego_, which sunk in 1600 AD [@Dizon2016; @Hsieh1995]. Southeast China is assumed to be the origin of the light gray glazed jars, however these are commonly found at sites in Taiwan that were associated with European activities, such as the Zeelandia fort site in Tainan [@Wang2007]. The jar shapes found at Kiwulan are typical of those found elsewhere in VOC sites occupied during the 17th century [@Berrocal2018, pp.917; @Cort2017, pp.282; @Grave2013; @Ketel2011; @Klose2018, pp.131]. We cannot be sure of the exact process that brought them to Kiwulan: they might have been directly imported by Europeans, by Chinese merchants, or by Indigenous groups via regional networks in north Taiwan. In any case, the high volume of ceramics transported by Europeans, and their high mobility in the shipping trade played an important role in introducing foreign jars to Taiwan.
